---
title: "Covid19 vaccine study Fall2021"
author: "Kayleen Sepulveda"
date: "December 1, 2021"
output: 
  html_document:
    toc: true
    toc_depth: 5

---





```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalData <-read.csv(file="fall21.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="guy")
gal = filter(FinalData, sex =="gal")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```



# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


$H_0$: There is no difference in covid-19 vaccine effectiveness among different subgroups(males,females,Drug users, LGBTQ)

$H_A$: there is a difference in covid-19 vaccine effectiveness among different subgroups(males,females,drug users, LGBTQ)


# Males

```{r}
dd2 <- guy %>% group_by(infected,treatment) %>%
summarize(count=n()) %>% mutate(prcnt=count/sum(count))


 # the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))

basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")
```




This plot represents the population of guys(males) people and the relationship to the contraction of covid-19 by either talking the vaccine or the placebo. 

The plot above shows that a shocking 58% of the people that took the vaccine did in fact get covid-19, while only approximately 42% of the people taking the placebo contracted covid-19. A substantially higher percentage of of people who took the vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be ineffective in protection against covid-19. The odds of getting covid-19 while taking the vaccine is 3:2. The men that took the drug were one and half times more likely to get covid-19 than those taking the placebo. 


There are two plots. in first plot, the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not. In the second plot, the one above the left two bars of the graph are all the people that got covid-19. the x-axis represents whether a person got covid-19 or not. the y-axis however, instead of representing the percentage of guys that got covid-19 after taking the placebo or taking the vaccine. 

The codes that create this plot start with the code block beggining with dd2. dd2 can be seen in the enviroment pane and if inspected contains 4 variables. these variables include infected, treatment, count, and percent. The second group of codes is called the basicC. BasicC can also be seen in the enviroment pane and appears to contain a list of 9 variables. This information creates the first plot. 


### Numerical results 


```{r}
table1 <- xtabs(~infected + treatment, data= guy)
rowPerc(table1)
colPerc(table1)
```
These numerical results show specific percentages of men that got covid-19 after taking the placebo and also taking the vaccine. 
Among the men 58% got covid-19 after taking the vaccine. while approximately 42% of the men that were given the placebo contracted covid-19. The vaccine is clearly not effective in preventing covid-19 in men. 

The odds of getting covid-19 while taking the vaccine is 3:2. The men that took the drug were one and a half times more likely to get covid-19 than those taking the placebo. 

### Inferntial results 


```{r}
chisq.test(table1)
chisqtestGC(table1)
guyfish <- fisher.test(table1)
guyfish
```
This information contains the results of the output of the Fisher exact test. The P-value displayed above is extremely small at 0.0005801. This number is very small, moreover seeing these resukts most likely concludes that the results are not likely to occur by chance. 

The null hypothesis predicted that the deaths related to covid-19 are independent of Subgroup. It is clear that Subgroups do play a role in the covid-19 infectionrate of the individuals, The alternative hypothesis , that there is a correlation between covid-19 vaccine effectiveness and subgroups is the favoirable option. 

The odd ratio above illustrates a ratio of 0.68 the men who recived the vaccine were 1.5 times more likely to get covid-19 than those who recieved the placebo. 

```{r}
effect <- guyfish$p.value/(guyfish$p.value +1)
effectivenessguy = guyfish$estimate/(guyfish$estimate +1)*100
print(effectivenessguy)
```

```{r}
low=(guyfish$conf.int[1]/(guyfish$conf.int[1] +1))*100
hi=(guyfish$conf.int[2]/(guyfish$conf.int[2] +1))*100
options(digits = 2)
```

above is the effectiveness for the covid-19 vaccine for males. if the effectiveness is above 50% that means that the vaccine is effective. based on the interval seen above, we can conclude that the vaccine is ineffective in protecting males from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for men can be seen above as 40. The effectiveness confidence interval for men is 35.09 to 45.94. 


# Females 



```{r}
dd2 <- gal %>% group_by(infected,treatment) %>%
summarize(count=n()) %>% mutate(prcnt=count/sum(count))


 # the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))

basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")
```

this plot represents the population of girls(females) and the relationship to the contraction of covid-19 by either taking the vaccine or the placebo. 

The plot above shows that only 33% of the girls that took the vaccine did in fact get covid-19. while 67% of the people taking the placebo contracted covid-19 .substantially hugher percentage of people who took vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be effective in protection against covid-19 for females. The odds of getting covid-19 while taking the vaccine is 1:3. The women that took the drug were half as likely to get covid-19 than those taking the placebo. 


There are two plots. in first plot, the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not. In the second plot, the one above the left two bars of the graph are all the people that got covid-19. the x-axis represents whether a person got covid-19 or not. the y-axis however, instead of representing the percentage of females that got covid-19 after taking the placebo or taking the vaccine. 

The codes that create this plot start with the code block beggining with dd2. dd2 can be seen in the enviroment pane and if inspected contains 4 variables. these variables include infected, treatment, count, and percent. The second group of codes is called the basicC. BasicC can also be seen in the enviroment pane and appears to contain a list of 9 variables. This information creates the first plot. 


### Numerical results 


```{r}
table1 <- xtabs(~infected + treatment, data= gal)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of females that got covid-19 after taking the placebo and taking the vaccine. 
Among the gals, 33% got covid-19 after taking the vaccine, while approximately 67% of the females that were given the placebo contracted covid-19. The vaccine is clearly effective in preventing covid-19 in females and it is not reccomended. Those who took the placebo were much less likely to contract covid-19. 
The odds of getting covid-19 for women while taking the vaccine is 1:2. the women that took the drug were half as likely to get covid-19(33%) than those taking the placebo(67%)


### Inferential results 


```{r}
chisq.test(table1)
chisqtestGC(table1)
galfish <- fisher.test(table1)
galfish
```


This information is important because among all of these results is the output of the fisher exact test. The p-value displayed above is extremely small at 2e-16. This is an awfully small number. This number means that these results are not likely to occur by chance. 

The null hypothesis predicted that the infection rate related to covid-19 vaccination effectiveness are independent of subgroup. Since it is clear that subgroup does play a role in the Covid infection rate in the vaccinated individuals, The alternative hypothesis is that there is a correlation between covid-19 vaccination effectiveness in subgroups.

The odd ratio above illustrates a ratio of 2.00. the females who recived the vaccine were half as likely to get covid-19 than those who recived the placebo. 





```{r}
effect <- galfish$p.value/(galfish$p.value +1)
effectivenessgal = galfish$estimate/(galfish$estimate +1)*100
print(effectivenessgal)
```

```{r}
low=(galfish$conf.int[1]/(galfish$conf.int[1] +1))*100
hi=(galfish$conf.int[2]/(galfish$conf.int[2] +1))*100
options(digits = 2)
```

above is the effectiveness for the covid-19 vaccine for females. if the effectiveness is above 50% that means that the vaccine is effective. based on the interval seen above, we can conclude that the vaccine is effective in protecting females from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for females can be seen above as 67. The effectiveness confidence interval for females is 55.09 to 75.05.



# Drug Users


```{r}
dd2 <- Druguser %>% group_by(infected,treatment) %>%
summarize(count=n()) %>% mutate(prcnt=count/sum(count))


 # the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))

basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")
```


this plot represents the population of drug users and the relationship to the contraction of covid-19 by either taking the vaccine or the placebo. 

The plot above shows that only 33% of the girls that took the vaccine did in fact get covid-19. while 67% of the drug users taking the placebo contracted covid-19 .substantially lower percentage of people who took vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be effective in protection against covid-19 for drug users. The odds of getting covid-19 while taking the vaccine is 1:3. The drug users that took the drug were half as likely to get covid-19 than those taking the placebo. 


There are two plots. in first plot, the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not. In the second plot, the one above the left two bars of the graph are all the drug users that got covid-19. the x-axis represents whether a person got covid-19 or not. the y-axis however, instead of representing the percentage of drug users that got covid-19 after taking the placebo or taking the vaccine. 

The codes that create this plot start with the code block beggining with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables. these variables include infected, treatment, count, and percent. The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. helps create the percentage plot This information creates the first plot. 


### Numerical results 


```{r}
table1 <- xtabs(~infected + treatment, data= Druguser)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of drug users that got covid-19 after taking the placebo and taking the vaccine. 
Among the drug users, 33% got covid-19 after taking the vaccine, while approximately 67% of the drug users that were given the placebo contracted covid-19. The vaccine is clearly effective in preventing covid-19 in drug users and it is not reccomended. Those who took the placebo were much more likely to contract covid-19. 
The odds of getting covid-19 for drug users while taking the vaccine is 1:2. the drug users that took the drug were half as likely to get covid-19 (33%) than those taking the placebo (67%). 


### Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
Druguserfish <- fisher.test(table1)
Druguserfish
```


This information is important because among all of these results is the output of the fisher exact test. The p-value displayed above is extremely small at .002 This is an awfully small number. This number means that these results are not likely to occur by chance. 

The null hypothesis predicted that the covid-19 vaccine effectiveness is independent of subgroup. Since it is clear that subgroup does play a role in the Covid infection rate in the vaccinated individuals, The alternative hypothesis is that there is a correlation between covid-19 vaccination effectiveness in subgroups.

The odd ratio above illustrates a ratio of 1.8 the drug users who received the vaccine were half as likely to get covid-19 than those who received the placebo. 


```{r}
effect <- Druguserfish$p.value/(Druguserfish$p.value +1)
effectivenessDruguser = Druguserfish$estimate/(Druguserfish$estimate +1)*100
print(effectivenessDruguser)
```

```{r}
low=(Druguserfish$conf.int[1]/(Druguserfish$conf.int[1] +1))*100
hi=(Druguserfish$conf.int[2]/(Druguserfish$conf.int[2] +1))*100
options(digits = 2)
```

above is the effectiveness for the covid-19 vaccine for drug users. if the effectiveness is above 50% that means that the vaccine is effective. based on the interval seen above, we can conclude that the vaccine is effective in protecting drug users from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for drug users can be seen above as 65. The effectiveness confidence interval for Drug users is 55.09 to 75.05.

# LGBTQ

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>%
summarize(count=n()) %>% mutate(prcnt=count/sum(count))


 # the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))

basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")
```

this plot represents the population of LGBTQ and the relationship to the contraction of covid-19 by either taking the vaccine or the placebo. 

The plot above shows that only 33% of the girls that took the vaccine did in fact get covid-19. while 67% of the drug users taking the placebo contracted covid-19 .substantially lower percentage of people who took vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be effective in protection against covid-19 for drug users. The odds of getting covid-19 while taking the vaccine is 1:3. The drug users that took the drug were half as likely to get covid-19 than those taking the placebo. 


There are two plots. in first plot, the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not. In the second plot, the one above the left two bars of the graph are all the drug users that got covid-19. the x-axis represents whether a person got covid-19 or not. the y-axis however, instead of representing the percentage of drug users that got covid-19 after taking the placebo or taking the vaccine. 

The codes that create this plot start with the code block beggining with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables. these variables include infected, treatment, count, and percent. The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. helps create the percentage plot This information creates the first plot. 


### Numerical results 

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of LGBTQ that got covid-19 after taking the placebo and taking the vaccine. 
Among the LGBTQ, 71% got covid-19 after taking the vaccine, while approximately 29% of the LGBTQ that were given the placebo contracted covid-19. The vaccine is clearly ineffective in preventing covid-19 in LGBTQ and it is not reccomended for them. Those who took the placebo were much more likely to contract covid-19. 
The odds of getting covid-19 for while taking the vaccine is 3:1 for LGBTQ. the LGBTQ that took the drug were much more likely to get covid-19 (71%) than those taking the placebo (29%). 


### Inferential results


```{r}
chisq.test(table1)
chisqtestGC(table1)
gayfish <- fisher.test(table1)
gayfish
```

This information is important because among all of these results is the output of the fisher exact test. The p-value displayed above is extremely small at .00006 This is an awfully small number. This number means that these results are not likely to occur by chance. 

The null hypothesis predicted that the covid-19 vaccine effectiveness is independent of subgroup. Since it is clear that subgroup does play a role in the Covid infection rate in the vaccinated individuals, The alternative hypothesis is that there is a correlation between covid-19 vaccination effectiveness in subgroups.

The odd ratio above illustrates a ratio of 0.4 the LGBTQ who received the vaccine were as likely to get covid-19 than those who received the placebo. 


```{r}
effect <- gayfish$p.value/(gayfish$p.value +1)
effectivenessgay = gayfish$estimate/(gayfish$estimate +1)*100
print(effectivenessgay)
```

```{r}
low=(gayfish$conf.int[1]/(gayfish$conf.int[1] +1))*100
hi=(gayfish$conf.int[2]/(gayfish$conf.int[2] +1))*100
options(digits = 2)
```


above is the effectiveness for the covid-19 vaccine for LGBTQ. if the effectiveness is above 50% that means that the vaccine is effective. based on the interval seen above, we can conclude that the vaccine is ineffective in protecting LGBTQ from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for LGBTQ can be seen above as 29. The effectiveness confidence interval for LGBTQ is 24.45 to 36.56



# Conclusions and recommendations

We reject the Null hypothesis because there is a difference in Covid-19 vaccine effectiveness among different subgroups(males,females,Drug users,LGBTQ). Vaccine effectiveness varies between subgroups and each subgroup had a different P-value, thus the discrepancy rate of effectiveness varies differently between subgroups after receiving the vaccine. 

Males vaccinated had 2:1 of contracting Covid-19 after vaccination, thus it is ineffective for males with a p-value of 0.0005801 which is too low. This indicates strong evidence against the null hypothesis, so we reject the null hypothesis. In conclusion males should not take the vaccine because it is ineffective at protecting males against covid-19. 

Females were 1:2 odds of contracting Covid-19 after vaccination with P-value of <2e-16

