---
title: "Covid19 vaccine study Fall2021"
author: "Put your name here!"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5

---





```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalData <-read.csv(file="fall21.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="guy")
gal = filter(FinalData, sex =="gal")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```



# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


$H_0$: There is no difference in covid-19 vaccine effectiveness among different subgroups(males,females,Drug users, LGBTQ)


# Males

```{r}
dd2 <- guy %>% group_by(infected,treatment) %>%
summarize(count=n()) %>% mutate(prcnt=count/sum(count))


 # the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))

basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")
```


This plot represents the population of guys(males) people and the relationship to the contraction of covid-19 by either talking the vaccine or the placebo. 

The plot above shows that a shocking 58% of the people that took the vaccine did in fact get covid-19, while only approximately 42% of the people taking the placebo contracted covid-19. A substantially higher percentage of of people who took the vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be ineffective in protection against covid-19. The odds of getting covid-19 while taking the vaccine is 3:2. The men that took the drug were one and half times more likely to get covid-19 than those taking the placebo. 


There are two plots. in first plot, the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not. In the second plot, the one above the left two bars of the graph are all the people that got covid-19. the x-axis represents whether a person got covid-19 or not. the y-axis however, instead of representing the percentage of guys that got covid-19 after taking the placebo or taking the vaccine. 

The codes that create this plot start with the code block beggining with dd2. dd2 can be seen in the enviroment pane and if inspected contains 4 variables. these variables include infected, treatment, count, and percent. The second 


# Females

# Drug Users

# LGBTQ

# Conclusions and recommendations

